### 2021-10-24 paper/pldi22/formal.tex
Till semantics
More formalization for function specialization

  \item \underline{Function Calls}:
    \[
      \sem{\varx = \expr \kwrl \expr_1 \cdots \expr_n \kwrr}(\st) =
      (\lab_\varf, \ctxt :: \ctxts, \heap, \env')
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \sem{\expr} &=& (\st_0, \kwdef \; \varf \; \kwrl \varp_1, \cdots,
        \varp_n \kwrr \; \lab_\varf) &\wedge\\
        \sem{\expr_k}(\st_{k-1}) &=& (\st_k, \val_k) \; \forall 1 \leq k \leq n
        &\wedge\\
        \st_n &=& (\lab, \ctxts, \heap, \env) &\wedge\\
        \env' &=& [\varp_1 \mapsto \val_1, \cdots, \varp_n \mapsto \val_n]
        &\wedge\\
        \ctxt &=& (\getnext(\lab), \env, \varx)\\
      \end{array}
    \]

@@

\paragraph{Expressions: $\framebox{$\sem{\expr}: \stset \rightarrow \stset
\times \valset$}$}
\begin{itemize}
  \item \underline{Primive Values}:
    \[
      \sem{\pval}(\st) = \todo
    \]
  \item \underline{Operations}:
    \[
      \sem{\op \kwrl \expr^* \kwrr}(\st) = \todo
    \]
  \item \underline{Object Allocations}:
    \[
      \sem{\kwcl \kwcr}(\st) = \todo
    \]
  \item \underline{References}:
    \[
      \sem{\refer}(\st) = \todo
    \]
\end{itemize}

\paragraph{References: $\framebox{$\sem{\refer}: \stset \rightarrow \stset
\times \valset$}$}
\begin{itemize}
  \item \underline{Variable Lookups}:
    \[
      \sem{\varx}(\st) = \todo
    \]
  \item \underline{Field Lookups}:
    \[
      \sem{\refer \kwsl \expr \kwsr}(\st) = \todo
    \]
\end{itemize}

@@

A function specialization is defined with the following three components:
\begin{itemize}
  \item Restriction $\restrictset$
  \item Concretization $\gamma: \restrictset \rightarrow \powerset{\stset
    \times \valset^*}$
  \item Transformation $\transform: \restrictset \rightarrow \funcset
    \rightarrow \funcset$.
\end{itemize}

\subsubsection{Constant Propagation Specialization}
\begin{itemize}
  \item Restriction $\cprestrictset = (\valoptset)^* =
    (\valset \uplus \{ \top \})^*$
  \item Concretization $\cpgamma: (\valoptset)^* \rightarrow \powerset{\stset
    \times \valset^*}$ s.t.
    \[
      \cpgamma(\valopt_1, \cdots, \valopt_n) = \stset \times (V_1 \times \cdots
      \times V_n)
    \]
    where $V_i = \left\{
      \begin{array}{ll}
        \valset & \text{if} \; \valopt_i = \top\\
        \{ \valopt_i \} & \text{otherwise}\\
      \end{array}
    \right.$
  \item Transformation $\cptransform: (\valoptset)^* \rightarrow \funcset
    \rightarrow \funcset$
\end{itemize}
In the constant propagation specialization, a restriction is a sequences of
static values $\valset$ or a dynamic value $\top$.

>>>>

  \item \underline{Function Calls}:
    \[
      \sem{\varx = \expr \kwrl \expr_1 \cdots \expr_n \kwrr}(\st) =
      (\lab_\varf, \ctxt :: \ctxts, \heap, \env')
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \sem{\expr}(\st) &=& (\st_0, \kwdef \; \varf \; \kwrl \varp_1, \cdots,
        \varp_n \kwrr \; \lab_\varf) &\wedge\\
        \sem{\expr_k}(\st_{k-1}) &=& (\st_k, \val_k) \; \forall 1 \leq k \leq n
        &\wedge\\
        \st_n &=& (\lab, \ctxts, \heap, \env) &\wedge\\
        \env' &=& [\varp_1 \mapsto \val_1, \cdots, \varp_n \mapsto \val_n]
        &\wedge\\
        \ctxt &=& (\getnext(\lab), \env, \varx)\\
      \end{array}
    \]

@@

\paragraph{Expressions: $\framebox{$\sem{\expr}: \stset \rightarrow \stset
\times \valset$}$}
\begin{itemize}
  \item \underline{Primive Values}:
    \[
      \sem{\pval}(\st) =
      (\st, \pval)
    \]
  \item \underline{Operations}:
    \[
      \sem{\op \kwrl \expr_1, \cdots, \expr_n \kwrr}(\st) =
      (\st_n, \op(\val_1, \cdots, \val_n))
    \]
    where
    \[
      \st_0 = \st \wedge
      \forall 1 \leq k \leq n. \; \sem{\expr_k}(\st_{k-1}) = (\st_k, \val_k)
    \]
  \item \underline{Object Allocations}:
    \[
      \sem{\kwcl \kwcr}(\st) =
      ((\lab, \ctxts, \heap[\addr \mapsto \varnothing], \env), \addr)
    \]
    where
    \[
      \st = (\lab, \ctxts, \heap, \env) \wedge
      \addr \not\in \text{Domain}(\heap)
    \]
\end{itemize}

\paragraph{References: $\framebox{$\sem{\refer}: \stset \rightarrow \stset
\times \valset$}$}
\begin{itemize}
  \item \underline{Variable Lookups}:
    \[
      \sem{\varx}(\st) =
      (\st, \env(\varx))
    \]
    where
    \[
      \st = (\_, \_, \_, \env)
    \]
  \item \underline{Field Lookups}:
    \[
      \sem{\refer \kwsl \expr \kwsr}(\st) =
      (\st'', \val')
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \sem{\refer}(\st) &=& (\st', \addr) &\wedge\\
        \sem{\expr}(\st') &=& (\st'', \str) &\wedge\\
        \st'' &=& (\_, \_, \heap, \_) &\wedge\\
        \obj &=& \heap(\addr) &\wedge\\
        \val' &=& \obj(\str)\\
      \end{array}
    \]
\end{itemize}

@@

A \textit{function specialization} is defined with the following three
components:
\begin{itemize}
  \item Restriction $\restrictset$
  \item Concretization $\gamma: \restrictset \rightarrow \powerset{\stset
    \times \valset^*}$
  \item Transformation $\transform: \restrictset \rightarrow \funcset
    \rightarrow \funcset$.
\end{itemize}
We can restrict possible input states and arguments of functions by defining a
set of restrictions $\restrictset$ with their concretization $\gamma:
\restrictset \rightarrow \powerset{\stset \times \valset^*}$. A transformation
$\transform: \restrictset \rightarrow \funcset \rightarrow \funcset$ is
parameteric with such restrictions and transforms a function to another function
by preserving observational semantics under a given restriction.

\subsubsection{Constant Propagation Specialization}
We define a \textit{constant propagation specialization} with the following
three components:

\begin{itemize}
  \item Restriction $\cprestrictset = (\valoptset)^* =
    (\valset \uplus \{ \top \})^*$
  \item Concretization $\cpgamma: (\valoptset)^* \rightarrow \powerset{\stset
    \times \valset^*}$ s.t.
    \[
      \cpgamma(\valopt_1, \cdots, \valopt_n) = \stset \times (V_1 \times \cdots
      \times V_n)
    \]
    where $V_i = \left\{
      \begin{array}{ll}
        \valset & \text{if} \; \valopt_i = \top\\
        \{ \valopt_i \} & \text{otherwise}\\
      \end{array}
    \right.$
  \item Transformation $\cptransform: (\valoptset)^* \rightarrow \funcset
    \rightarrow \funcset$
\end{itemize}
The constant propagation specialization restricts specific arguments with
concrete values with a restriction defined with a sequence of static or dynamic
values $\valset \uplus \{ \top \}$, but it does not restrict input states.
While a dynamic value $\top$ denotes no restriction, it restricts a specific
argument with the static value $\val \in \valset$. To specialize functions under
a given restriction $(\valopt_1, \cdots, \valopt_n) \in (\valoptset)^*$, we
define a transformation $\cptransform: (\valoptset)^* \rightarrow \funcset
\rightarrow \funcset$. To propagate constants, \todo What is a constant?
